---
title: "PCA on USArrest Data"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    number_sections: no
    fig_caption: true
    toc: true
    toc_float: true
    toc_depth: 3
    # code_folding: hide
header-includes:
- \usepackage{graphicx}
- \usepackage{float}
editor_options:
  chunk_output_type: inline
---

```{r setup knit, include=FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
	fig.pos = "H",
	message = FALSE,
	warning = FALSE,
	external = TRUE,
	echo = TRUE
)
```

```{r}
# import packages
library(tidyverse)   
library(magrittr)    
library(DT)         
library(kableExtra)

library(factoextra)  
```

# PCA on USArrests

### Preparing Data


*USArrests* data is a set that contains four variables that represent the number of arrests per 100,000 residents for *Murder*, *Assault* and *Rape* in each of the fifty US states in 1973. The data set also contains the percentage of the population living in urban areas, *UrbanPop*.

```{r}
# check data
datatable(USArrests, options = list(pageLength = 5))
```

```{r}
# scale data
scaled_df <- 
  apply(USArrests, 2, scale) %>% 
  set_colnames(paste0(names(USArrests), "_scaled")) %>% 
  set_rownames(rownames(USArrests))

datatable(round(scaled_df, 4), options = list(pageLength = 5))
```


### Eigenvectors & Eigenvalues



```{r}
# compute covariance matrix of scaled data 
covariance_matrix <- cov(scaled_df) 
# compute eigenvalues and eigenvector (or pca loadings)
Eigen <- eigen(covariance_matrix)
Eigen
```



### PC Scores



```{r}
# compute pc scores (data matrix multiples pca loadings)
pc_df <- 
  as.matrix(scaled_df) %*% Eigen$vectors %>% 
  data.frame() %>% 
  set_colnames(paste0("PC",1:ncol(scaled_df)))

datatable(round(pc_df, 4), options = list(pageLength = 5))
```



```{r}
# compare to results of built-in function
pca_result <- prcomp(USArrests, scale = TRUE)   # it can standardize data for us
datatable(round(pca_result$x, 4), options = list(pageLength = 5)) 
```


### Biplot


```{r}
fviz_pca_biplot(
  pca_result, 
  axes = c(1, 2),              # choose two PC's to plot
  geom = c("point", "text"),   # plot states as point, text labels, or both
  labelsize = 3,               # label font size 
  pointsize = 0.7,             # point size for states
  col.ind = "grey50",          # point and label color for states
  col.var = "contrib",         # use variable contributions to PC as legend 
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")    # color for variable contrib
  ) +
  xlim(c(-3.1,3.1)) + ylim(c(-2.5,2.5))
```


```{r}
# rotate 180 degree
fviz_pca_biplot(
  pca_result, 
  axes = c(1, 2),              
  geom = c("point", "text"),  
  labelsize = 3,               
  pointsize = 0.7,             
  col.ind = "grey50",          
  col.var = "contrib",         
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")    
  ) +
  xlim(c(3.1,-3.1)) + ylim(c(2.5,-2.5))      # rotate the biplot
```


# PCR on Imcome per Capita

### Extract income data

```{r}
library(rvest)

# URL of website that has data
url = 'https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_income#cite_note-4'

income_data = url %>%
  # Read the html page
  read_html() %>%  
  # Inspect the page, then find html code line that highlights table we are interested in, right-click and copy xpath
  html_node(xpath = '//*[@id="mw-content-text"]/div[1]/table[2]') %>%
  html_table(fill = TRUE) %>% 
  mutate(income = str_remove_all(`2018`,'[$,]') %>% as.numeric(),
         state = `State or territory`) %>% 
  select(state, income) %>% 
  left_join(
    USArrests %>% rownames_to_column('state'),
    .,
    by = 'state'
  )

datatable(income_data)
```

### Collinearity

```{r}
state = income_data$state
income_data = income_data[,-1]

corrplot::corrplot.mixed(cor(income_data))
```

### Cross Validation

```{r}
library(pls)
pcr_fit = pcr(income~., data = income_data, scale = TRUE, validation = "CV")
summary(pcr_fit)
```

### Validation plots {.tabset}

#### RMSEP

```{r}
validationplot(pcr_fit, val.type="RMSEP", cex.axis=0.7)
axis(side = 1, at = c(2), cex.axis=0.7)
abline(v = 2, col = "blue", lty = 3)
```

#### MSEP

```{r}
validationplot(pcr_fit, val.type="MSEP", cex.axis=0.7)
axis(side = 1, at = c(2), cex.axis=0.7)
abline(v = 2, col = "blue", lty = 3)
```


#### R2

```{r}
validationplot(pcr_fit, val.type="R2", cex.axis=0.7)
axis(side = 1, at = c(2), cex.axis=0.7)
abline(v = 2, col = "blue", lty = 3)
```

### Coefficients

```{r}
pcr_fit2 <- pcr(income~., data = income_data, scale = TRUE, ncomp = 2)

pcr_fit2$coefficient %>% 
  data.frame() %>% 
  round(2) %>% 
  set_colnames(paste(1:2,'comps')) %>% 
  kable() %>% 
  kable_styling()
```





