---
title: "PCR on Fullerton Housing"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    number_sections: no
    fig_caption: true
    toc: true
    toc_float: true
    toc_depth: 3
header-includes:
- \usepackage{graphicx}
- \usepackage{float}
editor_options:
  chunk_output_type: inline
---

```{r setup knit, include=FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
	fig.pos = "H",
	message = FALSE,
	warning = FALSE,
	external = TRUE,
	echo = TRUE
)

library(tidyverse)
library(magrittr)

library(DataExplorer)
library(lubridate)

library(pls)

library(ggsci)
library(kableExtra)
```


```{r}
# clean data 
Fullerton <- 
  read.csv('FullertonHousing.csv') %>% 
  mutate(SOLD_DATE = mdy(SOLD_DATE),
         # SOLD_MONTH = month(SOLD_DATE, label = TRUE),
         ZIP = as.factor(ZIP),
         LOT_SIZE = ifelse(is.na(LOT_SIZE), SQUARE_FEET, LOT_SIZE),
         HOA = ifelse(is.na(HOA), 0, HOA)) %>% 
  select(-c(SOLD_DATE,CITY,STATE,HOA))

str(Fullerton)
```

## Cross Validation

```{r}
set.seed(2)
pcr_fit <- pcr(PRICE~., data = Fullerton, scale = TRUE, validation = "CV")
summary(pcr_fit)
```

## Validation Plots {.tabset}

### RMSEP

```{r}
validationplot(pcr_fit, val.type="RMSEP", cex.axis=0.7)
axis(side = 1, at = c(7), cex.axis=0.7)
abline(v = 7, col = "blue", lty = 3)
```

### MSEP

```{r}
validationplot(pcr_fit, val.type="MSEP", cex.axis=0.7)
axis(side = 1, at = c(7), cex.axis=0.7)
abline(v = 7, col = "blue", lty = 3)
```

### R2

```{r}
validationplot(pcr_fit, val.type="R2", cex.axis=0.7)
axis(side = 1, at = c(7), cex.axis=0.7)
abline(v = 7, col = "blue", lty = 3)
```


## Coefficients

```{r}
pcr_fit2 <- pcr(PRICE~., data = Fullerton, scale = TRUE, ncomp = 7)

coef_tb <- pcr_fit2$coefficients %>% data.frame() 
coef_tb[,2:7] %>% 
  round(2) %>% 
  set_colnames(paste(c(2:7),'comps')) %>% 
  kable() %>% 
  kable_styling()
```


## ZIP 92831

### Cross Validation

```{r}
Fullerton_subset <- Fullerton %>% filter(ZIP == 92831) %>% select(-ZIP)

set.seed(2)
pcr_fit <- pcr(PRICE~., data = Fullerton_subset, scale = TRUE, validation = "CV")
summary(pcr_fit)
```

### Validation Plots {.tabset}

### RMSEP

```{r}
validationplot(pcr_fit, val.type="RMSEP", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### MSEP

```{r}
validationplot(pcr_fit, val.type="MSEP", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### R2

```{r}
validationplot(pcr_fit, val.type="R2", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### Coefficients

```{r}
pcr_fit2 <- pcr(PRICE~., data = Fullerton_subset, scale = TRUE, ncomp = 5)

coef_tb <- pcr_fit2$coefficients %>% data.frame() 
coef_tb[,2:5] %>% 
  round(2) %>% 
  set_colnames(paste(c(2:5),'comps')) %>% 
  kable() %>% 
  kable_styling()
```


## ZIP 92832

### Cross Validation

```{r}
Fullerton_subset <- Fullerton %>% filter(ZIP == 92832) %>% select(-ZIP)

set.seed(2)
pcr_fit <- pcr(PRICE~., data = Fullerton_subset, scale = TRUE, validation = "CV")
summary(pcr_fit)
```

### Validation Plots {.tabset}

### RMSEP

```{r}
validationplot(pcr_fit, val.type="RMSEP", cex.axis=0.7)
axis(side = 1, at = c(4), cex.axis=0.7)
abline(v = 4, col = "blue", lty = 3)
```


### MSEP

```{r}
validationplot(pcr_fit, val.type="MSEP", cex.axis=0.7)
axis(side = 1, at = c(4), cex.axis=0.7)
abline(v = 4, col = "blue", lty = 3)
```


### R2

```{r}
validationplot(pcr_fit, val.type="R2", cex.axis=0.7)
axis(side = 1, at = c(4), cex.axis=0.7)
abline(v = 4, col = "blue", lty = 3)
```


### Coefficients

```{r}
pcr_fit2 <- pcr(PRICE~., data = Fullerton_subset, scale = TRUE, ncomp = 4)

coef_tb <- pcr_fit2$coefficients %>% data.frame() 
coef_tb[,2:4] %>% 
  round(2) %>% 
  set_colnames(paste(c(2:4),'comps')) %>% 
  kable() %>% 
  kable_styling()
```


## ZIP 92833

### Cross Validation

```{r}
# droplevels: only two types of properties in 92833, either condo or single family residental
Fullerton_subset <- Fullerton %>% filter(ZIP == 92833) %>% select(-ZIP) %>% droplevels() 

set.seed(2)
pcr_fit <- pcr(PRICE~., data = Fullerton_subset, scale = TRUE, validation = "CV")
summary(pcr_fit)
```

### Validation Plots {.tabset}

### RMSEP

```{r}
validationplot(pcr_fit, val.type="RMSEP", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### MSEP

```{r}
validationplot(pcr_fit, val.type="MSEP", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### R2

```{r}
validationplot(pcr_fit, val.type="R2", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### Coefficients

```{r}
pcr_fit2 <- pcr(PRICE~., data = Fullerton_subset, scale = TRUE, ncomp = 5)

coef_tb <- pcr_fit2$coefficients %>% data.frame() 
coef_tb[,2:5] %>% 
  round(2) %>% 
  set_colnames(paste(c(2:5),'comps')) %>% 
  kable() %>% 
  kable_styling()
```


## ZIP 92835

### Cross Validation

```{r}
Fullerton_subset <- Fullerton %>% filter(ZIP == 92835) %>% select(-ZIP)

set.seed(2)
pcr_fit <- pcr(PRICE~., data = Fullerton_subset, scale = TRUE, validation = "CV")
summary(pcr_fit)
```

### Validation Plots {.tabset}

### RMSEP

```{r}
validationplot(pcr_fit, val.type="RMSEP", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### MSEP

```{r}
validationplot(pcr_fit, val.type="MSEP", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### R2

```{r}
validationplot(pcr_fit, val.type="R2", cex.axis=0.7)
axis(side = 1, at = c(5), cex.axis=0.7)
abline(v = 5, col = "blue", lty = 3)
```


### Coefficients

```{r}
pcr_fit2 <- pcr(PRICE~., data = Fullerton_subset, scale = TRUE, ncomp = 5)

coef_tb <- pcr_fit2$coefficients %>% data.frame() 
coef_tb[,2:5] %>% 
  round(2) %>% 
  set_colnames(paste(c(2:5),'comps')) %>% 
  kable() %>% 
  kable_styling()
```




